---
title: "Building a CV from tidy data storage"
author: "Eli Felts"
format: html
editor: visual
bibliography: references.bib
---

```{r options echo=FALSE}

knitr::opts_chunk$set(
  message=FALSE,
  warning=FALSE,
  fig.width=6.5,
  fig.height=6,
  cache = TRUE,
  dev="png"
)

library(tidyverse)
library(vitae)
library(conflicted)
library(RefManageR)
library(googlesheets4)
library(googledrive)

conflicts_prefer(dplyr::filter(),
                 dplyr::lag())
```

# Initial Steps

To generate PDF documents from RMarkdown you will need to setup something called LaTeX on your local machine. You can do so using the `tinytex` package [@tinytex] with the following code.

```{r eval=FALSE}

tinytex::install_tinytex()

```

You will only have to do this once. Now, to use templates from `vitae` make sure to install that package as well. You can do so from the RStudio menu by clicking Tools -\> Install Packages and searching for `vitae`, or simply run the following code.

```{r}

install.packages("vitae")

```

# Starting from template

To start off, open a new RMarkdown file and on the left of the dialog box you will have an option to select "From Template". This will show templates that are available from the `vitae` package. To view examples of these templates check out the [README](https://github.com/mitchelloharawild/vitae?tab=readme-ov-file){target="_blank"} for the package.
